---
title: "Week 10 - Day 2 - Homework"
output: html_notebook
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

<hr>


### This is unfinished, I was mostly playing around to see what different code/
models would do. I understand regressions etc but think I have some confusion 
between models and the "long way" of doing things and have been trying to do all 
at once. 



```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
library(mosaic)
```

# MVP

<br>

1. Load the `housing_prices.csv` data set and undertake an initial exploration 
of the data. You will find details on the data set on the relevant [Kaggle page](https://www.kaggle.com/camnugent/california-housing-prices)

```{r}
housing_prices <- read_csv("data/housing_prices.csv")

summary(housing_prices)
# 200 NAs in total_bedrooms. ALl numeric apart from ocean_proximity (chr)
```


<br>

2. We expect the `total_rooms` of houses to be strongly correlated with 
`total_bedrooms`. Use `ggpairs()` to investigate correlations between these 
two variables.

```{r}
housing_prices %>% 
  select(total_rooms, total_bedrooms) %>% 
  ggpairs()

# very strong and positive correlation (r = 0.934). There is a cluster in the
# lower values and plots lessen as both rooms and bedrooms increase
```


<br> 

3. So, we do find significant correlations. Let's drop `total_bedrooms` from the 
dataset, and use only `total_rooms` going forward.

```{r}
housing_prices <- housing_prices %>% 
  select(-total_bedrooms)
```


<br>

4. We are interested in developing a regression model for the `median_house_value` 
of a house in terms of the possible predictor variables in the dataset. 

  i. Use `ggpairs()` to investigate correlations between `median_house_value` 
  and the predictors (this may take a while to run, don't worry, make coffee or 
  something).
  
```{r}
# predictors: housing_median_age (maybe income is better but I still
# think it could be interesting), total_rooms, population, households (maybe take
# out this or population, wait and see), median_income, ocean_proximity (is that
# enough or is lat and long needed too - join these)

housing_prices %>% 
  ggpairs()

housing_prices
```


  ii. Perform further `ggplot` visualisations of any significant correlations 
  you find.

```{r}
# Population isn't significant but others are - is ocean_proximity?!
# median income and total rooms have the strongest correlations, so does
# latitude but i'd want to consider that joined to longitude

median_value_interest <- housing_prices %>% 
  select(- population)

median_value_interest %>% 
  ggplot()+
  aes(x = median_house_value, 
      y = median_income)+
  geom_point() +
  geom_smooth()

median_value_interest %>% 
  ggplot()+
  aes(x = median_house_value, 
      y = total_rooms)+
  geom_point()+
  geom_smooth()
```


<br>

5. Shortly we may try a regression model to fit the categorical predictor 
`ocean_proximity`. Investigate the level of `ocean_proximity` predictors. How 
many dummy variables do you expect to get from it?
    
```{r}
housing_prices %>% 
  distinct(ocean_proximity)

# 5 dummy variables 
```


<br>
 
6. Start with simple linear regression. Regress `median_house_value` on 
`median_income` and check the regression diagnostics.

```{r}
# dependent = house value as that is what is being studied
model <- lm(formula = median_house_value ~ median_income, 
            data = housing_prices)

summary(model)
# intercept = 45457, for every increase of 1 value in 

library(modelr)
library(ggfortify) # need for autoplot

autoplot(model)

```


<br> 

7. Add another predictor of your choice. Check your assumptions, diagnostics, 
and interpret the model.

```{r}
model2 <- lm(formula = median_house_value ~ median_income + ocean_proximity, 
            data = housing_prices)

summary(model2)

# <1hr from the ocean is our 
# fewer inland houses than <1hr from ocean but more of others (especially island)
```


# Extension
    
<br>

8. Try adding an interaction between `log(medium_income)` and your chosen categorical predictor. Do you think this interaction term is statistically justified?

<br>

9. Find and plot an appropriate visualisation to show the effect of this interaction


