---
title: "Week 6 - Weekend Homework - Mahri"
output: html_notebook
---

# Hypothesis tests & probability - homework

## MVP

### Hypothesis testing - practical

You work for an animal conservation organisation and are looking to do some 
investigating into some of the animals to provide evidence for some projects 
you are looking to run. 

In this homework we'll use the `msleep` data set provided in the `ggplot` 
package. It concerns the sleeping patterns of various types of mammal.

```{r}
library(tidyverse)
library(ggplot2)
library(janitor)
library(infer)
```

<br>

**Question 1.**  
Explore the dataset and familiarise yourself with it.

```{r}
data(msleep)
glimpse(msleep)
head(msleep, 10)

```

<br>

**Question 2.**  
Jabberwockies sleep for around 7 hours a night, on average. Perform an 
appropriate statistical test to determine whether the mean `sleep_total` in the 
sampled population of animal types differs from the typical value for 
jabberwockies.

```{r}
obs_stat_sleep_total <- msleep %>% 
  summarise(mean_sleep = round(mean(sleep_total), 2))
obs_stat_sleep_total
# The mean sleep total for all animals = 10.43 hours

# Does the mean DIFFER from the population sample mean - two tailed as it's not
# asking if it's more or less.

# H0: There is no difference between the mean hours of sleep for jabberwockies
# and the population.
# H1: The mean hours of sleep for the population differs from that of the 
# jabberwockies.

# create a null distribution
# 7 is the average hours of sleep for the jabberwockie (mu = hypothesis mean
#  and we're hypothesizing that the population mean is the same)
null_distribution_total_sleep <- msleep %>% 
  specify(response = sleep_total) %>% 
  hypothesise(null = "point", mu = 7) %>% 
  generate(reps = 50000, type = "bootstrap") %>% 
  calculate(stat = "mean")
null_distribution_total_sleep
 

# Visualise. The red line will be our observed stat i.e. the mean for the 
# population.
# The bars will be the means of means.
null_distribution_total_sleep %>% 
  visualise(bins = 30) +
  shade_p_value(obs_stat = obs_stat_sleep_total, direction = "both")

# p value 
p_value_sleep_total <- null_distribution_total_sleep %>% 
  get_p_value(obs_stat = obs_stat_sleep_total$mean_sleep, direction = "both")
p_value_sleep_total

# p = 0.
# As p < 0.05, we can reject our null hypothesis in favour of the alternative
# hypotheseis - the mean hours of sleep for the population differs from that of 
# the jabberwockies.
```


<br>

**Question 3.**  
Perform an appropriate statistical test to determine whether omnivores sleep for 
significantly longer than herbivores, on average.

```{r}
# This is one tailed (right-tailed) ("do they sleep MORE")

# H0: Omnivores do not sleep for a significantly longer time than herbivores. 
# H1: Omnivores sleep for a significantly longer time than herbivores. 


omnis_and_herbis <- msleep %>% 
  filter(vore %in% c("omni", "herbi"))

omnis_and_herbis %>% 
  group_by(vore) %>% 
  summarise(count = n())
# There are 20 omnivores and 32 Herbivores


omnis_and_herbis %>% 
  ggplot()+ 
  aes(x = sleep_total, y = vore) +
  geom_boxplot() 
# Just having a look. The average scores are quite close but very different 
# ranges - omnivores mostly sleep for similar amounts of time with some 
# outliers, herbivores much more spread 
```

Theoretical distribution: (observed below)
```{r}
null_distribution_omni_herbi <- omnis_and_herbis %>% 
  specify(sleep_total ~ vore) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 5000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))
null_distribution_omni_herbi
```

Observed/ empirical distribution (theoretical above) and p-value:
```{r}
observed_stat_omni_herbi <- omnis_and_herbis %>% 
  specify(sleep_total ~ vore) %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))
observed_stat_omni_herbi
# 1.42 (it's positive (just?) so shows that omnis sleep for longer than herbis
# (?)


null_distribution_omni_herbi %>% 
  get_p_value(obs_stat = observed_stat_omni_herbi, direction = "right")

# As p > 0.05 (p = 0.12) we can not reject the null hypothesis that omnivores
# do not sleep for a significantly longer time than herbivores. 
```
Visualise and p-value:
One tail so shading only on the right. Red line = observed stat
```{r}
null_distribution_omni_herbi %>% 
  visualise() + 
  shade_p_value(obs_stat = observed_stat_omni_herbi, direction = "right")


```



<br>

**Question 4. **
Perform an appropriate statistical test to determine whether the proportion of 
domesticated animal types in the population of animal types is greater than 5%
**Hint**
Think about creating an `is_domesticated` variable for the analysis


```{r}
# one tailed test (domesticated greater than 5%) and PROPORTION not mean

# H0: Domesticated animals do not make up more than 5% of the population 
# H1: Domesticated animals make up at least 5% of the population

# Create is_domesticated column 
# There were: True = 10, False = 44, NA = 29
# Should I take out the NAs? Having looked at them, I'd say it's very unlikely 
# many of them have been domesticated, so I haven't... 
domesticated_or_not <- msleep %>% 
  mutate(conservation = replace_na(conservation, "no")) %>% 
  mutate(is_domesticated = conservation == "domesticated", .after = 5)
domesticated_or_not

domesticated_or_not %>% 
  group_by(is_domesticated) %>% 
  summarise(count = n())
# Now 73 false and 10 true.

# Looking at the proportion of domesticated animals
domesticated_or_not %>%
  group_by(is_domesticated) %>%
  summarise(prop = n()/nrow(domesticated_or_not)) %>%
  filter(is_domesticated == TRUE)
# 0.12 = 12% which is of course higher than 5% but we want to do a statistical 
# test to make sure this is not due to sampling variation. (this is the observed
# stat using dplyr which I've done again below with infer functions)
```

Drawing: 
```{r}
null_distribution_domesticated <- domesticated_or_not %>%
  specify(response = is_domesticated, success = "TRUE") %>% 
  hypothesize(null = "point", p = 0.05) %>%  # this 0.05 is because we're interested in 5%
  generate(reps = 5000, type = "draw") %>%
  calculate(stat = "prop") #specify that the stat is proportion 
null_distribution_domesticated
```

visualise: 
```{r}
null_distribution_domesticated %>% 
  visualise(bins = 30)
```

observed stat (calculated above with dplyr but now using infer functions for use
in hypothesis testing): 
```{r}
observed_stat_domesticated <- domesticated_or_not %>% 
  specify(response = is_domesticated, success = "TRUE") %>% 
  calculate(stat = "prop")

observed_stat_domesticated
# 0.12 
```
```{r}
null_distribution_domesticated %>%
  visualise(bins = 30) + #remembering that it is a '+' here not '%>%' because using ggplot functionality to visualise
  shade_p_value(obs_stat = observed_stat_domesticated, direction = "both")
```
p value:
```{r}
p_value_domesticated <- null_distribution_domesticated %>%
  get_p_value(obs_stat = observed_stat_domesticated, direction = "both")

p_value_domesticated

# p = 0.02 (0.0156) 
# As p < 0.05 we can reject the null hypothesis in favour of the alternative 
# hypothesis that domesticated animals make up at least 5% of the population

# Please note I changed NAs to FALSE so the homework answer might be different
```



<hr>

## Hypothesis testing - Defining the Hypothesis 

For the following three business problems write out:

* What kind of test you would use  
*  $H_0$ and $H_a$ in **both** mathematical notation and in words. 
* Also state the method you would use to generate the null distribution 
(bootstrap, permutation or simulation). 

<br>

**Question 1.**  
You work for a independent coffee shop. The boss tells you that she thinks that 
around $40\%$ of people in the town (population $30,000$) know of the coffee 
shop, but you are skeptical of the estimate. You conduct a random survey of 
$200$ people in the town, asking if respondents are aware of your coffee shop. 
You want to use the sample to test the hypothesis that $40\%$ or more of the 
town's population have heard of the coffee shop. 

<br>

**Question 2.**  
You work for a website design company and have performed an **A/B test** on the 
position of a banner on a website promoting a particular item. 
<br><br>
<div class='emphasis'>
**A/B testing** 
A method comparing two versions of a web page, email, flyer or other marketing 
device against each other to determine which version performs better. As it is 
essentially a controlled experiment, the design should try to ensure that the
groups experiencing both versions of the marketing device are equivalent and 
representative of the population.
</div>
<br><br>
You selected five days at random last month and then randomly selected $200$ of
each sampled day's users into group $A$ and another $200$ of that day's users 
into group $B$. Group $A$ continued to be shown the banner at the right hand 
side of the webpage (its usual position) while group $B$ was shown the banner 
at the top of the page. You monitored each sampled user's interactions with the 
website, in particular the 'click through rate' (CTR) on the banner, i.e. what 
proportion of sampled users clicked on the banner. You want to use the sampled 
data to test the hypothesis that website users overall are more likely to click 
on the banner if positioned at the top of the page   
<br>

**Question 3.**  
You work as an analyst for a car manufacturing company - they have specific 
standards they must meet for standards and regulation purposes. You have been 
asked to check the quality control of the manufacture of a particular car part. 
You have been given data on a sample of $200$ parts produced over the period of 
a week (the sampled parts were pulled at random from the production line and 
measured: the factory produced a lot more than $200$ parts that week). The 
specification of the part stipulates a width of $145\textrm{mm}$, and the 
manufacturing process is said to have 'drifted' if the mean width of parts 
differs significantly from $145\textrm{mm}$. You want to use the sampled
measurements to test whether the process overall has drifted.     




## Hypothesis Testing - Interpreting the results

For the 3 business problems stated above, imagine we performed you got the 
following p-values (with the given significance levels) write out your 
interpretation of the results. 

<br>

**Question 1.**  

**Coffee shop problem**. Significance level: 0.05, calculated $p$-value: 0.07

<br>

**Question 2.**  

**Website company problem**. Significance level: 0.01, $p$-value: 0.006

<br>

**Question 3.**  

**Manufacturing company problem**. Significance level: 0.05, $p$-value: 0.55

