---
title: "Final SQL Lab and Homework - Solutions"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r, eval=TRUE, include=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(RPostgres)

db = dbConnect(
  Postgres(), 
  user = 'omni_user',
  password = 'niytna5mcwsz124i',
  dbname = 'omni_pool',
  host = 'db-postgresql-lon1-20135-do-user-7581050-0.b.db.ondigitalocean.com',
  port = 25061,
  sslmode = 'require',
  bigint = 'numeric'
)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```
  
**Duration - 3 hours**  
  
In these practice questions you'll return again to the `omni_pool` database, so re-establish your connection in `DBeaver` if it has lapsed.

<br>
<div class='emphasis'>
**Submission**

When answering these questions, please save your `SQL` script in `DBeaver`, copy it over to your homework repo and then add, commit and push that for submission tonight. Please add section and question numbers as `SQL` comments, like so:

```{sql, eval=FALSE}
--MVP
--Q1
```
</div>
<br>

# MVP Questions

<br>

**Question 1.**  
How many employee records are lacking both a grade and salary?


```{sql, connection = db, output.var="out"}
SELECT
  count(id)
FROM employees
WHERE grade IS NULL AND salary IS NULL
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 2.**  
Produce a table with the two following fields (columns):

- the department
- the employees full name (first and last name)

Order your resulting table alphabetically by department, and then by last name

```{sql, connection = db, output.var="out"}
SELECT
  CONCAT(first_name, ' ', last_name) AS name, 
  department
FROM employees
ORDER BY department ASC NULLS LAST, last_name ASC NULLS LAST
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 3.**  
Find the details of the top ten highest paid employees who have a last_name beginning with 'A'.

```{sql, connection = db, output.var="out"}
SELECT
  *
FROM employees
WHERE last_name LIKE 'A%'
ORDER BY SALARY DESC NULLS LAST
LIMIT 10
```
```{r, echo=FALSE, eval=TRUE}
out
```


<br>


**Question 4.**  
Obtain a count by `department` of the employees who started work with the corporation in 2003.  

```{sql, connection = db, output.var="out"}
SELECT 
  department, 
  COUNT(id) AS num_employees_started_2003
FROM employees
WHERE start_date BETWEEN '2003-01-01' AND '2003-12-31'
GROUP BY department
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 5.**  
Obtain a table showing `department`, `fte_hours` and the number of employees in each department who work each `fte_hours` pattern. Order the table alphabetically by `department`, and then in ascending order of `fte_hours`. 
<br>
<details>
<summary>**Hint**</summary>
You need to `GROUP BY` two columns here.
</details>

```{sql, connection = db, output.var="out"}
SELECT
  department,
  fte_hours,
  COUNT(id) AS num_employees
FROM employees 
GROUP BY department, fte_hours
ORDER BY department ASC NULLS LAST, fte_hours ASC NULLS LAST
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 6.**  
Provide a breakdown of the numbers of employees enrolled, not enrolled, and with unknown enrollment status in the corporation pension scheme.

```{sql, connection = db, output.var="out"}
SELECT 
  pension_enrol AS pension_enrolled, 
  COUNT(id) AS num_employees
FROM employees
GROUP BY pension_enrol
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 7.**  
Obtain the details for the employee with the highest salary in the ‘Accounting’ department who is not enrolled in the pension scheme?

```{sql, connection = db, output.var="out"}
SELECT *
FROM employees
WHERE department = 'Accounting' AND pension_enrol IS FALSE
ORDER BY salary DESC NULLS LAST
LIMIT 1
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 8.**  
Get a table of `country`, number of employees in that country, and the average `salary` of employees in that country for any countries in which more than 30 employees are based. Order the table by average `salary` descending.  
<br>
<details>
<summary>**Hints**</summary>  
A `HAVING` clause is needed to filter using an aggregate function.
<br><br>
You can pass a column alias to `ORDER BY`.
</details>

```{sql, connection = db, output.var="out"}
SELECT 
  country, 
  count(id) AS num_employees, 
  AVG(salary) AS average_salary
FROM employees
GROUP BY country
HAVING COUNT(id) > 30
ORDER BY average_salary DESC
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 9.**   
11. Return a table containing each employees `first_name`, `last_name`, full-time equivalent hours (`fte_hours`), `salary`, and a new column `effective_yearly_salary` which should contain `fte_hours` multiplied by `salary`. Return only rows where `effective_yearly_salary` is more than 30000.

```{sql, connection = db, output.var="out"}
SELECT 
  first_name,
  last_name,
  fte_hours,
  salary,
  fte_hours * salary AS effective_yearly_salary
FROM employees
WHERE fte_hours * salary > 30000
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>


**Question 10**  
Find the details of all employees in either Data Team 1 or Data Team 2
<br>
<details>
<summary>**Hint**</summary>  
`name` is a field in table `teams
</details>

```{sql, connection = db, output.var = "out"}
SELECT *
FROM employees AS e INNER JOIN
  teams AS t ON e.team_id = t.id
WHERE t.name LIKE 'Data Team%'
```
```{r, echo=FALSE, eval=TRUE}
out
```



**Question 11**  
Find the first name and last name of all employees who lack a `local_tax_code`.  
<br>
<details>
<summary>**Hint**</summary>  
`local_tax_code` is a field in table `pay_details`, and `first_name` and `last_name` are fields in table employees
</details>

```{sql, connection = db, output.var = "out"}
SELECT 
  first_name, 
  last_name
FROM employees AS e INNER JOIN pay_details AS pd
ON e.pay_detail_id = pd.id
WHERE pd.local_tax_code IS NULL
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 12.**  
The `expected_profit` of an employee is defined as (48 * 35 * `charge_cost` - `salary`) * `fte_hours`, where `charge_cost` depends upon the team to which the employee belongs. Get a table showing `expected_profit` for each employee.  
<br>
<details>
<summary>**Hints**</summary> 
`charge_cost` is in `teams`, while `salary` and `fte_hours` are in `employees`, so a join will be necessary
<br><br>
You will need to change the type of `charge_cost` in order to perform the calculation
</details>

```{sql, connection = db, output.var = "out"}
SELECT 
  e.id,
  e.first_name,
  e.last_name,
  e.fte_hours,
  e.salary,
  t.name,
  t.charge_cost,
  (48 * 35 * t.charge_cost::INT - e.salary) * e.fte_hours
    AS expected_profit
FROM employees AS e LEFT JOIN teams AS t
ON e.team_id = t.id
```
```{r, echo=FALSE, eval=TRUE}
out
```   

<br>

**Question 13. [Tough]**  
Find the first_name, last_name and salary of the lowest paid employee in Japan
who works the least common full-time equivalent hours across the corporation.”

<details>
<summary>**Hint**</summary>
You will need to use a subquery to calculate the mode
</details>


```{sql, connection = db, output.var="out"}
SELECT
  first_name,
  last_name,
  salary
FROM employees
WHERE country = 'Japan' AND fte_hours = (
  SELECT fte_hours
  FROM employees
  GROUP BY fte_hours
  ORDER BY COUNT(*) DESC NULLS LAST
  LIMIT 1
  )
ORDER BY salary ASC NULLS LAST
LIMIT 1
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 14.**  
Obtain a table showing any `department`s in which there are two or more employees lacking a stored first name. Order the table in descending order of the number of employees lacking a first name, and then in alphabetical order by `department`.

```{sql, connection = db, output.var="out"}
SELECT department, COUNT(id) AS num_employees_no_first
FROM employees 
WHERE first_name IS NULL
GROUP BY department
HAVING COUNT(id) >= 2
ORDER BY COUNT(id) DESC NULLS LAST, department ASC NULLS LAST 
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 15. [Bit tougher]**  
Return a table of those employee `first_name`s shared by more than one employee, together with a count of the number of times each `first_name` occurs. Omit employees without a stored `first_name` from the table. Order the table descending by count, and then alphabetically by `first_name`.

```{sql, connection = db, output.var="out"}
SELECT 
  first_name, 
  COUNT(id) AS name_count
FROM employees
WHERE first_name IS NOT NULL
GROUP BY first_name 
HAVING COUNT(id) > 1
ORDER BY COUNT(id) DESC, first_name ASC
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>


**Question 16. [Tough]**  
Find the proportion of employees in each department who are `grade` 1.
<br>
<details>
<summary>**Hints**</summary>
Think of the desired proportion for a given `department` as the number of employees in that department who are `grade` 1, divided by the total number of employees in that department.  
<br><br>
You can write an expression in a `SELECT` statement, e.g. `grade = 1`. This would result in `BOOLEAN` values.  
<br><br>
If you could convert `BOOLEAN` to `INTEGER` 1 and 0, you could sum them. The `CAST()` function lets you convert data types.  
<br><br>
In `SQL`, an `INTEGER` divided by an `INTEGER` yields an `INTEGER`. To get a `REAL` value, you need to convert the top, bottom or both sides of the division to `REAL`.
</details>

```{sql, connection = db, output.var="out"}
SELECT 
  department, 
  SUM(CAST(grade = '1' AS INT)) / CAST(COUNT(id) AS REAL) AS prop_grade_1 
FROM employees 
GROUP BY department
```
```{r, echo=FALSE, eval=TRUE}
out
```

or in PostgreSQL, can write type conversion as:

```{sql, connection = db, output.var="out"}
SELECT
  department, 
  SUM((grade = '1')::INT) / COUNT(id)::REAL AS prop_grade_1 
FROM employees 
GROUP BY department
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

# Extension Questions

Some of these problems may need you to do some online research on `SQL` statements we haven't seen in the lessons up until now... Don't worry, we'll give you pointers, and it's good practice looking up help and answers online, data analysts and programmers do this all the time! If you get stuck, it might also help to sketch out a rough version of the table you want on paper (the column headings and first few rows are usually enough). 

Note that some of these questions may be best answered using CTEs or window functions: have a look at the optional lesson included in today's notes!

**Question 1. [Tough]**  
Get a list of the `id`, `first_name`, `last_name`, `department`, `salary` and `fte_hours` of employees in the **largest** department. Add two extra columns showing the ratio of each employee's `salary` to that department's average `salary`, and each employee's `fte_hours` to that department's average `fte_hours`.  
<br>
[**Extension - really tough!** - how could you generalise your query to be able to handle the fact that two or more departments may be **tied** in their counts of employees. In that case, we probably don't want to arbitrarily return details for employees in just one of these departments].
<br>
<details>
<summary>**Hints**</summary> 
Writing a CTE to calculate the name, average `salary` and average `fte_hours` of the largest department is an efficient way to do this.
<br><br>
Another solution might involve combining a subquery with window functions
<br>
</details>

**CTE solution**

```{sql, connection = db, output.var = "out"}
WITH biggest_dept_details(name, avg_salary, avg_fte_hours) AS (
  SELECT 
     department,
     AVG(salary),
     AVG(fte_hours)
  FROM employees
  GROUP BY department
  ORDER BY COUNT(id) DESC NULLS LAST
  LIMIT 1
)
SELECT
  e.id,
  e.first_name,
  e.last_name,
  e.department,
  e.salary,
  e.fte_hours,
  e.salary / bdd.avg_salary AS salary_over_dept_avg,
  e.fte_hours / bdd.avg_fte_hours AS fte_hours_over_dept_avg
FROM employees AS e INNER JOIN biggest_dept_details AS bdd
ON  e.department = bdd.name
```
```{r, echo=FALSE, eval=TRUE}
out
```

**Window function solution**

```{sql, connection = db, output.var = "out"}
SELECT 
	id, 
	first_name, 
	last_name, 
	department,
	salary,
	fte_hours,
	salary / AVG(salary) OVER () AS salary_over_dept_avg,
	fte_hours / AVG(fte_hours) OVER () AS fte_hours_over_dept_avg
FROM employees
WHERE department = (
  SELECT
    department
  FROM employees
  GROUP BY department
  ORDER BY COUNT(id) DESC NULLS LAST
  LIMIT 1
);
```
```{r, echo=FALSE, eval=TRUE}
out
```

The problem with using `ORDER BY... LIMIT 1` to find the largest or smallest entry in a table is that there may be a **tie**, and `LIMIT 1` would then arbitrarily return just one of the tied rows. We might want to code to account for this possibility with something like this:

**CTE solution for ties**

```{sql, connection = db, output.var = "out"}
WITH all_dept_details(name, count, avg_salary, avg_fte_hours) AS (
  SELECT
    department,
    COUNT(id),
    AVG(salary),
    AVG(fte_hours)
  FROM employees
  GROUP BY department
), 
biggest_dept_count(max_count) AS (
  SELECT
    MAX(count)
  FROM all_dept_details
),
biggest_dept_details AS (
  SELECT 
    name,
    avg_salary,
    avg_fte_hours
  FROM all_dept_details INNER JOIN biggest_dept_count
  ON all_dept_details.count = biggest_dept_count.max_count
)
SELECT
  e.id,
  e.first_name,
  e.last_name, 
  e.department,
  e.salary,
  e.fte_hours,
  e.salary / bdd.avg_salary AS salary_over_dept_avg,
  e.fte_hours / bdd.avg_fte_hours AS fte_hours_over_dept_avg
FROM employees AS e INNER JOIN biggest_dept_details AS bdd
ON e.department = bdd.name
```
```{r, echo=FALSE, eval=TRUE}
out
```

**Window function solution for ties**

```{sql, connection = db, output.var = "out"}
SELECT 
	id, 
	first_name, 
	last_name, 
	department,
	salary,
	fte_hours,
	salary / AVG(salary) OVER (PARTITION BY department) AS salary_over_dept_avg,
	fte_hours / AVG(fte_hours) OVER (PARTITION BY department) AS fte_hours_over_dept_avg
FROM employees
WHERE department IN (
  SELECT
    department
  FROM employees
  GROUP BY department
  HAVING COUNT(id) = (
	  SELECT
		  MAX(count)
	  FROM (
		  SELECT
			  department,
			  COUNT(id)
		  FROM employees 
		  GROUP BY department
	  ) AS temp
  )
);
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 2.**  
Have a look again at your table for MVP question 6. It will likely contain a blank cell for the row relating to employees with 'unknown' pension enrollment status. This is ambiguous: it would be better if this cell contained 'unknown' or something similar. Can you find a way to do this, perhaps using a combination of `COALESCE()` and `CAST()`, or a `CASE` statement?  
<br>
<details>
<summary>**Hints**</summary>
`COALESCE()` lets you substitute a chosen value for `NULL`s in a column, e.g. `COALESCE(text_column, 'unknown')` would substitute `'unknown'` for every `NULL` in `text_column`. The substituted value has to match the **data type** of the column otherwise `PostgreSQL` will return an error.
<br><br>
`CAST()` let's you change the data type of a column, e.g. `CAST(boolean_column AS VARCHAR)` will turn `TRUE` values in `boolean_column` into text `'true'`, `FALSE` to text `'false'`, and will leave `NULL`s as `NULL`
</details>

**A `COALESCE()` and `CAST()` solution:**

```{sql, connection = db, output.var="out"}
SELECT 
  COALESCE(CAST(pension_enrol AS VARCHAR), 'unknown') AS pension_enrolled, 
  COUNT(id) AS num_employees
FROM employees
GROUP BY pension_enrol
```
```{r, echo=FALSE, eval=TRUE}
out
```

**A `CASE` solution:**

```{sql, connection = db, output.var="out"}
SELECT 
  CASE 
    WHEN pension_enrol IS NULL THEN 'unknown'
    WHEN pension_enrol IS TRUE THEN 'yes' 
    ELSE 'no'
  END AS pension_enrolled, 
  COUNT(id) AS num_employees
FROM employees
GROUP BY pension_enrol
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 3.**  Find the first name, last name, email address and start date of all the employees who are members of the 'Equality and Diversity' committee. Order the member employees by their length of service in the company, longest first.

```{sql, connection = db, output.var = "out"}
SELECT 
  e.first_name, 
  e.last_name, 
  e.email, 
  e.start_date
FROM 
employees AS e INNER JOIN employees_committees AS ec
ON e.id = ec.employee_id
INNER JOIN committees AS c
ON ec.committee_id = c.id
WHERE c.name = 'Equality and Diversity'
ORDER BY e.start_date ASC NULLS LAST
```
```{r, echo=FALSE, eval=TRUE}
out
```

<br>

**Question 4. [Tough!]**  
Use a `CASE()` operator to group employees who are members of committees into `salary_class` of `'low'` (`salary < 40000)` or `'high'` (`salary >= 40000`). A `NULL` `salary` should lead to `'none'` in `salary_class`. Count the number of committee members in each `salary_class`.  
<br>
<details>
<summary>**Hints**</summary>
You likely want to count `DISTINCT()` employees in each `salary_class`
<br><br>
You will need to `GROUP BY salary_class`
<br>
</details>

```{sql, connection = db, output.var = "out"}
SELECT 
  CASE 
    WHEN e.salary < 40000 THEN 'low'
    WHEN e.salary IS NULL THEN 'none'
    ELSE 'high' 
  END AS salary_class,
  COUNT(DISTINCT(e.id)) AS num_committee_members
FROM employees AS e INNER JOIN employees_committees AS ec
ON e.id = ec.employee_id
INNER JOIN committees AS c
ON ec.committee_id = c.id
GROUP BY salary_class
```
```{r, echo=FALSE, eval=TRUE}
out
```

```{r, echo=FALSE, eval=TRUE}
dbDisconnect(db)
```