---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
books <- read_csv("data/books.csv")
```
```{r}
view(books)
```

```{r}
dim(books) 
# there are 11,123 rows and 13 columns

names(books)

count(books, language_code) 
# there are 27 language codes (including en-CA, en-GB, en-US, and eng)

books %>%
  summarise(count = sum(is.na(books))) #there are no NAs... doesn't mean there aren't any 0's or -'s


glimpse(books) 
# remember there are character and numeric data types 

max(books, average_rating) #WHY DO I GET THIS ERROR: object 'average_rating' not found
```




Which are the top 5 languages rated here and how many are in variations of english language Vs other languages

```{r}

language_grouping_count <- books %>% 
  select(language_code) %>% 
  count(language_code) %>% 
  arrange(desc(n))
language_grouping_count

# there are 27 language codes:
# 1 - English - almost 8908
# 2 - English-US - 1408
# 3 - Spanish - 218
# 4 - English-GB - 214  (Just my interest - what is the diff between eng and this?)
# 5 - French - 144



filter(books, language_code %in% c("eng", "en-US", "en-GB", "en-CA"))

# 10,537 are in a variation of English (586 are therefore not)



non_english_language <- books %>% 
  filter(language_code != "eng" & 
           language_code != "en-US" & 
           language_code != "en-GB" & 
           language_code != "en-CA")
non_english_language

#yep, there are 586 of them... couldn't work out if there is a cleaner way of getting this info

```


Calculating the average rating across all books and creating a new column that states TRUE if the book has a rating > the average and FALSE if not (my column appeared... then disappeared!? HELP)

```{r}

mean_average_rating <- books %>% 
  mutate(mean_average_rating = mean(average_rating, na.rm = TRUE))
mean_average_rating
  
# average rating across all books is 3.93(4075) out of 5
#next step (deleted) didn't work but realised I need for the mean average rating to be a numeric rather than a tibble so have to do:

mean_average_rating <- books %>% 
  summarise(average_rating = mean(average_rating, na.rm = TRUE)) %>% 
  pull()
class(mean_average_rating)

# now mean_average_rating is a numeric 

books %>% 
  mutate(rating_above_average = average_rating > mean_average_rating)
books


#I can see below that it's worked and added the column... but then taken it away again so I can't see it (nothing in the Environment panel either)... (error -  no applicable method for 'mutate' applied to an object of class "c('double', 'numeric')" when added below pull())


# average rating across all books is 3.93(4075) out of 5
# (My interest - some books have a rating of 0, does this mean NA or does it mean they were actually rubbish?! I don't think NA because the author says it's well done)


```


There are lots with 0 pages / with very small numbers of pages... some are definitely children's books, so let's go with anything less than 10 should be removed.

```{r}

remove_short_books <- books %>% 
  select(title, average_rating, num_pages) %>% 
  mutate(more_than_10pages = num_pages >= 10) %>% 
  mutate(more_than_10pages = na_if(more_than_10pages, FALSE)) %>% 
  filter(!is.na(more_than_10pages))
remove_short_books

# There MUST be a better way...
# added a new column called "more than 10pages" with TRUE or FALSE as to whether there's >= 10 pages. Changed Falses to NAs. Removed those with less than 10. 
 # 10,928 rows remaining / 11,123


```

Using case_when to add a new column showing description of length


```{r}

#from the internet... doesn't work 
# pages_range <- books %>% 
#  select(num_pages) %>% 
#  group_by(grp = cut(value, c(-Inf, 2000, 4000, Inf))) %>%
#  summarise(count = sum(count))
# pages_range

short_or_long <- books %>% 
  select(num_pages) %>% 
  mutate(short_or_long = case_when(
    num_pages <= 50 ~ "Very Short",
    num_pages <= 300 ~ "Short",
    num_pages <= 1000 ~ "Medium",
    num_pages <= 2000 ~ "Long",
    TRUE ~ "Very Long"
  ))
short_or_long
  


```


Working out a link between length of book and ratings_count / text_reviews_count (though looking at both ends of the table, it doesn't really seem like there's a terribly distinct pattern)... not finished

```{r}

length_vs_count <- books %>% 
  select(title, num_pages, ratings_count, text_reviews_count)
length_vs_count

```




More stuff that didn't work - What is the mean rating average for english, all english, and foreign language books

```{r}

# Looking at the average rating for English books - (remember there are 8908/11123 of them)

eng_language_ave_rating <- books %>% 
  filter(language_code == "eng") %>% 
  mutate(eng_average_rating = mean(average_rating, na.rm = TRUE)) %>% 
eng_language_ave_rating

# average for eng (English) books is 3.93(4062) (unsurprisingly same as average for all)

# what is the average for all versions of English...

all_english_language <- books %>% 
  filter(language_code %in% c("eng", "en-US", "en-GB", "en-CA")) %>% 
  mutate(all_english_language, all_eng_average_rating = mean(average_rating, na.rm = TRUE)) %>% 
all_english_language

# I can filter for language code but then it doesn't let me mutate for average_rating 
# ERROR MESSAGE Error in all_english_language(.) : could not find function "all_english_language"
# I've tried assigning all english language too...


# I can do this: 
# all_english_language <- books %>% 
# filter(language_code %in% c("eng", "en-US", "en-GB", "en-CA")) %>% 
# all_english_language

# and get that 10537 language codes include English (586 therefore in another language)



# so i'll look at non-english language...
         
non_english_language <- books %>% 
  filter(language_code != "eng" & language_code != "en-US" & language_code != "en-GB" & language_code != "en-CA")
non_english_language

#yep, there are 586 of them as found before...

non_english_ave_rating <- non_english_language %>% 
  mutate(non_engl_average_rating = mean(average_rating, na.rm = TRUE)) %>% 
non_english_ave_rating

# Error in non_english_ave_rating(.) : could not find function "non_english_ave_rating" 
```







