---
title: "Clustering homework"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings. 

* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers. 

* Perform k-means clustering and chose a value of k.

* Visualise the clustering for your chosen value of k.

* Do you think the clustering seems a good fit for this data?

* Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson). 


```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
```

```{r}
mall_customers <- read_csv("data/mall_customers.csv")
```
annual income (k$)
Spending Score(1-100)

```{r}
glimpse(mall_customers)

mall_customers <- mall_customers %>% 
  janitor::clean_names()

summary(mall_customers)

mall_customers %>% 
  group_by(age) %>% 
  summarise(count = n()) 
#51 ages, 18 - 70 years.

mall_customers %>% 
  group_by(spending_score_1_100) %>% 
  summarise(count = n()) 
#84 spending scores, 1 - 99.

mall_customers %>% 
  group_by(gender) %>% 
  summarise(n = n())
# 112 female, 88 male. 
```

200 customers aged 18 - 70 years with a minimum annual income of $15k, and max
of $137k (mean = 60.56k). Spending scores range from 1-100 but the max is only 
99 (mean = 50.20). 112 are female and 88 are male. 


```{r}
mall_customers %>% 
  ggplot()+
  aes(x = age, 
      fill = gender)+ 
  geom_bar()
```

```{r}
mall_customers %>% 
  select(-customer_id) %>% 
ggpairs()
```
There is a significant (but quite low) negative correlation between `spending score` 
and `age` (-0.33).  

There are _potential_ clusters in the `annual income` and `spending score` graph, 
and in the `age` and `spending score` graph. 


```{r}
mall_customers %>% 
  select(annual_income_k, spending_score_1_100) %>% 
  ggplot(aes(x = annual_income_k, 
             y = spending_score_1_100))+
  geom_point()
```


...Should I scale?? Annual income is in the 1000s and spending score is less than
100. So, probably. 

```{r}
mall_customers %>% 
  as_tibble() %>% 
  pivot_longer(cols = c(annual_income_k, spending_score_1_100),
               names_to = "type", 
               values_to = "value") %>% 
  group_by(type) %>% 
  summarise(mean = round(mean(value)),
            sd = sd(value))

```
annual income mean = 61, sd = 26.26
spending score mean = 50, sd = 25.82


```{r}
score_and_income <- mall_customers %>% 
  select(spending_score_1_100, annual_income_k) 

score_and_income_scale <- score_and_income %>% 
  mutate_all(scale)


score_and_income_scale %>% 
  as_tibble() %>% 
  pivot_longer(cols = c(annual_income_k, spending_score_1_100),
               names_to = "type", 
               values_to = "value") %>% 
  group_by(type) %>% 
  summarise(mean = round(mean(value)),
            sd = sd(value))
```   



### How do we find out how many clusters will be best: 

#### 1. Elbow method

```{r}
library(factoextra)

fviz_nbclust(score_and_income_scale,
             kmeans, 
             method = "wss",
             nstart = 25
             )
```

From this graph, I might choose 3 clusters, or maybe 5 is sharper? I'll see 
what the other methods suggest but from looking at the scatter plot, I reckon 
5 makes sense. 


#### 2. Silhouette coefficient

```{r}
fviz_nbclust(score_and_income_scale, 
             kmeans, 
             method = "silhouette", 
             nstart = 25)
```

5 clusters is the optimum here, but 6 is pretty close. 


#### 3. Gap Statistics 

fviz_nbclust(students_big_scale, kmeans, method = "gap_stat") #would put nstart=25 if had more computing power

```{r}
fviz_nbclust(score_and_income_scale, 
             kmeans, 
             method = "gap_stat")
```

Oh, this is messy. It says 1. But is it really 3? 6?

Maybe the data is not well suited to k-means clustering as we have quite 
different outcomes from the three methods.



#### clustering with k-means

```{r}
set.seed(9) # remember only do this when playing about - don't do in real life

clustered_score_income <- kmeans(score_and_income_scale, 
                        centers = 5, # "5 centroids (clusters), please"
                        nstart = 25  # how many random sets would we like - runs 25 times for this model
                        )
clustered_score_income
```

5 clusters of sizes 39, 23, 35, 81, 22

^ That's quite a difference e.g. 81 and 22 (does it matter?)

^ For `within cluster sum of squares` (measure of error) the smaller the better
as it means clusters are quite tightly packed. Here, the figures range from 5.23
to 19.67, that's quite a difference (I think).


Looking at each cluster in a clearer way:
```{r}
library(broom)

tidy(clustered_score_income, 
     col.names = colnames(score_and_income_scale))
```


augment - looking at the full data set again:
```{r}
s_i_clusters <- augment(clustered_score_income, mall_customers)
s_i_clusters
```

"This is a handy way of running many models and saving all the results and/or 
performance metrics into a single tibble for more streamlined manipulation and 
comparison using ‘tidy’ principles."

```{r}
# Set min & max number of clusters want to look at 
max_k <- 5

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(score_and_income_scale, 
                             .x, 
                             nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, score_and_income)
  )

k_clusters
```

Visualising: 
```{r}
k_clusters %>% 
  unnest(cols = c(augmented)) %>% 
  ggplot(aes(x = annual_income_k, 
             y = spending_score_1_100)) +
  geom_point(aes(color = .cluster))
```


What are the average scores split by cluster?

```{r}
s_i_clusters %>% 
  group_by(.cluster) %>% 
  summarise(mean(annual_income_k), 
            mean(spending_score_1_100),
            mean(age))
```

Clusters:

1. 39 customers - Second highest annual income and high(est) spending score with 
an average age of ~33 years. 
Marketing should be looked into with this group - are they shopping for 
themselves/dependents?

2. 23 customers - Low annual income and low spending score with an average age 
of ~45 years. 

3. 35 customers - High(est) annual income and low(est) spending score with an 
average age of 41 years. Maybe they normally shop in high end shops that are not 
in this particular mall?

4. 81 customers - Medium annual income and medium spending score with an average
age of 43 years. I would think that looking into marketing towards these shoppers 
might be wise. They might not spend as much but there seem to be lots of them and
they might be buying on behalf of children/dependents as well as themselves. 

5. 22 customers - low(est) annual income but second highest spending score with 
an average age of 25 years. I would also consider looking into marketing here
especially due to their average age - are they students/ new to the workforce/
more likely to buy lots of outfits due to socialising/ weddings etc. Do they have
young children? etc. 

