---
title: "SQL Day 1 Homework - Solutions"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---


```{r, eval=TRUE, include=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(RPostgres)

db = dbConnect(
  Postgres(), 
  user = 'omni_user',
  password = 'niytna5mcwsz124i',
  dbname = 'omni_pool',
  host = 'db-postgresql-lon1-20135-do-user-7581050-0.b.db.ondigitalocean.com',
  port = 25061,
  sslmode = 'require',
  bigint = 'numeric'
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```
  
In these practice questions you'll return again to the `omni_pool` database, so re-establish your connection in `DBeaver` if it has lapsed. 

<br>
<div class='emphasis'>
**Submission**

When answering these questions, please save your `SQL` script in `DBeaver`, copy it over to your homework repo and then add, commit and push that for submission tonight. Please add section and question numbers as `SQL` comments, like so:

```{sql, eval=FALSE}
/* MVP */
/* Q1 */
```
</div>
<br>


# MVP

<br>

**Question 1.**\
Find all the employees who work in the 'Human Resources' `department`.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT * 
FROM employees
WHERE department = 'Human Resources'
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 2.**\
Get the `first_name`, `last_name`, and `country` of the employees who work in the 'Legal' `department`.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  first_name, 
  last_name, 
  country 
FROM employees
WHERE department = 'Legal'
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 3.**\
Count the number of employees based in Portugal.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  COUNT(id) AS num_employees_Portugal
FROM employees
WHERE country = 'Portugal'
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 4.**\
Count the number of employees based in either Portugal or Spain.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  COUNT(id) AS num_employees_Portugal_Spain
FROM employees
WHERE country IN ('Portugal', 'Spain')
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 5.**\
Count the number of pay_details records lacking a `local_account_no`.
<br><br>
<details>
<summary>**Hint**</summary>
Run a query first to get some records in the `pay_details` table, just to see what the data looks like
</details>

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  COUNT(id) AS num_pay_details_no_local_acct
FROM pay_details
WHERE local_account_no IS NULL
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 6.**\
Are there any pay_details records lacking both a `local_account_no` and `iban` number?

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  COUNT(id) AS num_pay_details_no_local_acct_iban
FROM pay_details
WHERE local_account_no IS NULL AND iban IS NULL
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 7.**\
Get a table with employees `first_name` and `last_name` ordered alphabetically by `last_name` (put any `NULL`s last).

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  first_name, 
  last_name
FROM employees
ORDER BY last_name ASC NULLS LAST
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 8.**  
Get a table of employees `first_name`, `last_name` and `country`, ordered alphabetically first by `country` and then by `last_name` (put any `NULL`s last).

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  first_name, 
  last_name,
  country
FROM employees
ORDER BY 
  country ASC NULLS LAST, 
  last_name ASC NULLS LAST
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 9.**  
Find the details of the top ten highest paid employees in the corporation.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT *
FROM employees
ORDER BY salary DESC NULLS LAST
LIMIT 10
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 10.**  
Find the `first_name`, `last_name` and `salary` of the lowest paid employee in Hungary.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  first_name, 
  last_name, 
  salary
FROM employees
WHERE country = 'Hungary'
ORDER BY salary ASC NULLS LAST
LIMIT 1
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 11.**\
How many employees have a `first_name` beginning with 'F'?

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  COUNT(id) AS num_employees_first_name_F
FROM employees
WHERE first_name ILIKE 'F%'
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 12.**  
Find all the details of any employees with a 'yahoo' `email` address?

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT *
FROM employees
WHERE email ILIKE '%yahoo%'
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>


**Question 13.**
Count the number of pension enrolled employees not based in either France or Germany.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  COUNT(id) AS num_pension_not_France_Germany
FROM employees
WHERE pension_enrol IS TRUE AND country NOT IN ('France', 'Germany')
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 14.**  
What is the maximum `salary` among those employees in the 'Engineering' department who work 1.0 full-time equivalent hours (`fte_hours`)?

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  MAX(salary) AS max_salary_engineering_ft
FROM employees
WHERE department = 'Engineering' AND fte_hours = 1
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 15.**  
 Return a table containing each employees `first_name`, `last_name`, full-time equivalent hours (`fte_hours`), `salary`, and a new column `effective_yearly_salary` which should contain `fte_hours` multiplied by `salary`.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  first_name,
  last_name,
  fte_hours,
  salary,
  fte_hours * salary AS effective_yearly_salary
FROM employees
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

# Extension

<br>

**Question 16.**  
The corporation wants to make name badges for a forthcoming conference. Return a column `badge_label` showing employees' `first_name` and `last_name` joined together with their `department` in the following style: 'Bob Smith - Legal'. Restrict output to only those employees with stored `first_name`, `last_name` and `department`. 

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT
  first_name,
  last_name,
  department,
  CONCAT(first_name, ' ', last_name, ' - ', department) AS badge_label
FROM employees
WHERE 
  first_name IS NOT NULL AND 
  last_name IS NOT NULL AND 
  department IS NOT NULL
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 17.**  
One of the conference organisers thinks it would be nice to add the year of the employees' `start_date` to the `badge_label` to celebrate long-standing colleagues, in the following style 'Bob Smith - Legal (joined 1998)'. Further restrict output to only those employees with a stored `start_date`.
<br><br>
[If you're really keen - try adding the month as a string: 'Bob Smith - Legal (joined July 1998)'] 
<br><br>
<details>
<summary>**Hints**</summary>
* Do some research on the `SQL` `EXTRACT()` function. 
* For the extension to the extension (!), the `PostgreSQL` `TO_CHAR()` function might help. 
* To be honest, date and time handling is one of the areas in which the functionality of specific RDBMSs deviates most notably from the `ANSI SQL` standard.
</details>
  
<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT
  first_name,
  last_name,
  department,
  start_date,
  CONCAT(
    first_name, ' ', last_name, ' - ', department, 
    ' (joined ', EXTRACT(YEAR FROM start_date), ')'
  ) AS badge_label
FROM employees
WHERE 
  first_name IS NOT NULL AND 
  last_name IS NOT NULL AND 
  department IS NOT NULL AND
  start_date IS NOT NULL
```
```{r, echo=FALSE, eval=TRUE}
out
```

```{sql, connection = db, output.var="out"}
SELECT
  first_name,
  last_name,
  department,
  start_date,
  CONCAT(
    first_name, ' ', last_name, ' - ', department, ' (joined ', 
    TO_CHAR(start_date, 'FMMonth'), ' ', TO_CHAR(start_date, 'YYYY'), ')'
  ) AS badge_label
FROM employees
WHERE 
  first_name IS NOT NULL AND 
  last_name IS NOT NULL AND 
  department IS NOT NULL AND
  start_date IS NOT NULL
```
```{r, echo=FALSE, eval=TRUE}
out
```
Note: In this last solution we added `FM` in front of `Month`. When using the `TO_CHAR()` function, its default is to add leading and trailing spaces to the string. `FM` removes these spaces.
</details>
<br>

**Question 18.**  
Return the `first_name`, `last_name` and `salary` of all employees together with a new column called `salary_class` with a value `'low'` where `salary` is less than 40,000 and value `'high'` where `salary` is greater than or equal to 40,000.
<br><br>
<details>
<summary>**Hints**</summary>
* Investigate how to use a `SQL` `CASE` statement to return the required values `'low'` and `'high'` based on the value of `salary`.
* Think carefully how to deal with `NULL` salaries.
</details>
  
<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  first_name, 
  last_name, 
  CASE 
    WHEN salary < 40000 THEN 'low'
    WHEN salary IS NULL THEN NULL
    ELSE 'high' 
  END AS salary_class
FROM employees
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
  
```{r}
dbDisconnect(db)
```


