---
title: "wk3_markdownHmwrk"
output: html_document
date: '2022-03-26'
---

**I'm not sure where I'm meant to do this work ... tables and the meme didn't work on "MARKDOWN Edit"**


```{r}
# Recreate the HTML document below using markdown
```

**Example**

<!DOCTYPE html>
<html lang="en">
<head>
</head>
<body>
  <h1>Welcome to our page!</h1>
  <p>This is an example of how we <em>could</em> lay something out.</p>
  <h3>Smaller headings are fun too!</h3>
  <ul>
    <li>This element is part of a list</li>
    <li>So is this one</li>
    <li><a href="https://http.cat/200">This one's a link too!</a></li>
  </ul>
</body>
</html> 


**MY RECREATION**

# Welcome to our page!

This is an example of how we _could_ lay something out. 

### Smaller headings are fun too!

* This element is part of a list 
* So is this one 
* [This one's a link too!](https://http.cat/200)


```{r}
# Task - 10 mins

#Use the table generator to make a table storing some information. Friends’ contact details, your classmates’ favourite colours, whatever you like!
#If you’re feeling a bit more adventurous, try importing a .csv file or building a table using one of the other markup languages
```




| Monday  	| Tuesday    	| Wednesday    	| Thursday   	| Friday            	|
|---------	|------------	|--------------	|------------	|-------------------	|
| Cycle   	| Spin Class 	| Arms and Abs 	| Lower Body 	| Full body stretch 	|
| Stretch 	| Pilates    	| Stretch      	| Run        	| Relax             	|
|         	|            	|              	| Stretch!   	|                   	|



```{r}
library(meme)
u <- "https://i.imgflip.com/ohrrn.jpg"
meme(u, "code", "all the things!")
```

