---
title: "CI and Bootstrap Homework - Week 6 - Day 3"
output: html_notebook
---

**1 MVP**
Now we’ll go back to CI creation in the normal fashion. We’ll take the ames data 
from the CIs lab earlier today and regard it now as a sample, we won’t be drawing 
any smaller samples from within it. This is the usual situation in an analysis: 
you use all the data available to you!

```{r}
library(tidyverse)
library(infer)
library(janitor)
library(here)
```

**Task 1.**
Load the data again, clean_names(), and re-familiarise yourself with it

```{r}
ames <- read_csv("data/ames.csv") %>% 
  clean_names()

glimpse(ames)
```


**Task 2.**
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?
```{r}
ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(fill = "sea green", col = "black", alpha = 0.7)+ 
  coord_cartesian(xlim = c(1000, 220000))


# lot_area is right skewed
# if I put col="white" in the geom_histogram, you can't really see the full
# extent of the right skew. 
```



**Task 3.**
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) 
of the sold houses.

```{r}
# we're not taking smaller samples so is the sample size 2930? (there are 2930 
# rows)
# my computer won't run 2930 happily, it keeps terminating. so I'll try with 
# 2000 (which took at least a minute):

bootstrap_sampling <- ames %>% 
  rep_sample_n(size = 2000, replace = TRUE, reps = 5000) %>% 
  summarise(
    mean_lot_area = mean(lot_area)
  )

bootstrap_sampling


bootstrap_sampling_hist <- bootstrap_sampling %>% 
  ggplot() +
  aes(x = mean_lot_area) + 
  geom_histogram(fill = "sea green", col = "black")+ 
  coord_cartesian(xlim = c(9500, 11000))
bootstrap_sampling_hist
```




**Task 4.**
Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution

```{r}
# ci_95 <- bootstrap_sampling %>% 
#   summarise(
#     mean = mean(mean_lot_area), 
#     lower_bound = quantile(mean_lot_area, probs = 0.025),
#     upper_bound = quantile(mean_lot_area, probs = 0.975)
#   )
# 
# ci_95
# Mean = 10,145.5
# Lower Bound = 9,830.748
# Upper Bound = 10,516.35



infer_resample <- ames %>% 
  #which variable we are interested in
  specify(response = lot_area) %>% 
  #generate the replicates to resample
  generate(reps = 5000, type = "bootstrap") %>% 
  # statistic we are interested in for each resample
  calculate(stat = "mean")

infer_resample

infer_ci_95 <- infer_resample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")

infer_ci_95
# lower ci = 9,871.338
# upper ci = 10,446.97


infer_resample %>% 
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_95)

```


Task 5.
You would like to know the mean(lot_area) of the sold houses with higher 
confidence. Calculate the 99% CI for this variable (you can re-use your 
bootstrap distribution from above). Is it narrower or broader than the 95% CI? 
Does that make sense?

```{r}
infer_resample <- ames %>% 
  #which variable we are interested in
  specify(response = lot_area) %>% 
  #generate the replicates to resample
  generate(reps = 5000, type = "bootstrap") %>% 
  # statistic we are interested in for each resample
  calculate(stat = "mean")

infer_resample

infer_ci_99 <- infer_resample %>% 
  get_confidence_interval(level = 0.99, type = "percentile")

infer_ci_99
# lower ci = 9,789.697
# upper ci = 10,523.41


infer_resample %>% 
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_99)

# The 99% CI is broader. This does make sense as the wider the sample (?) the 
# more we chances we give ourselves of having the mean included

```



Task 6.
Calculate the point estimate of the mean(lot_area)


```{r}

point_estimate_lot_area <- ames %>% 
  rep_sample_n(size = 2930, reps = 1) %>% 
  summarise(
    mean_lot_area = mean(lot_area)
  )
  
point_estimate_lot_area

# The mean for lot_area is 10,147.92
```



2 Extension


Task 1.
Calculate a point estimate and 95% CI for the proportion of houses in the data 
built before 1920. Does the number of reps you use matter? 
[Investigate reps from 200 up to 50000, memory of your laptop permitting].


oops... i thought i finished this!
```{r}
built_before_1920 <- ames %>% 
  filter(year_built <= 1920)
built_before_1920
#there are 233 of them

point_estimate_1920 <- built_before_1920 %>% 
  rep_sample_n(size = 233, reps = 200) %>% 
  summarise(
    mean_lot_area = mean(lot_area)
  )
  
point_estimate_1920

point_estimate_1920_more_reps <- built_before_1920 %>% 
  rep_sample_n(size = 233, reps = 1000) %>% 
  summarise(
    mean_lot_area = mean(lot_area)
  )
  
point_estimate_1920_more_reps
```

