---
title: "Logistic regression homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center")
```

# MVP

You have been provided with a set of data on customer purchases of either 'Citrus Hill' (`purchase = 'CH'`) or 'Minute Maid' (`purchase = 'MM'`) orange juice, together with some further attributes of both the customer and the store of purchase. A data dictionary is also provided in the `data` directory.

We would like you to build the best **predictive classifier** you can of whether a customer is likely to buy Citrus Hill or Minute Maid juice. Use **logistic regression** to do this. You should use either train-test splitting or cross-validation to evaluate your classifier. The metric for 'best classifier' will be **highest AUC value** either in the test set (for train-test splitting) or from cross-validation.

**Issues we faced, thoughts we had**

* This is quite a tough, open-ended exercise. We decided early on to use an automated approach to model selection using `glmulti()`, but feel free to use a manual approach if you prefer!
* The `Purchase` dependent variable will require wrangling to work in logistic regression. We replaced it with a `purchase_mm` logical variable.
* Wrangle other categorical variables to be factors too.
* `WeekOfPurchase` is also quite tough to deal with: should it be added as a factor variable (it will lead to many coefficients), left as numeric, or omitted entirely? See if you can come up with a strategy to decide what to do with it.
* Check for aliased variables and remove any aliases before you set off to find your best models. Remember, you can use something like `alias(purchase_mm ~ ., data = oj)` to do this, the dot `.` here means 'all variables'. Aliased variables will be listed down the left-hand side, and you can subsequently remove them.

**`glmulti()` hints**

If you decide to use `glmulti()` be prepared for your `R` session to hang if you decide to abort a run! The reason for this is that `glmulti()` actually uses a separate Java runtime to do its thing in the background, and unfortunately `R` can't instruct Java to terminate on request. D'oh! Accordingly, make sure you **save any changes** to your work **before** each `glmulti()` run. That way, you can force quit `RStudio` if necessary without losing work. 

Here are some example inputs for using `glmulti()` with logistic regression for a variety of purposes.
_see homework folder_

<hr>
<hr>

```{r}
library(tidyverse)

library(GGally)

```

```{r}
orange_juice <- read_csv("data/orange_juice.csv")
glimpse(orange_juice)
```

```{r}
# store 7 has it's own column... 
orange_juice %>% 
  group_by(StoreID) %>%     # there are 5 stores. 
  summarise(count = n()) # store 7 has 356, 2 has 222, 3 = 196, 1 = 157, 4 = 139
```
`store` = 0 if the `store_id` == 7, and is equal to it's store_id otherwise..

It's messy:
```{r}
orange_juice <- orange_juice %>% 
  janitor::clean_names() %>% 
  mutate(purchase = as_factor(purchase),
         #weekof_purchase = as_factor(weekof_purchase), #causes problems later
        #store7 = as.logical(store7 == "Yes"), # I tried lots but nothing worked
        store7 = as_factor(store7),
        special_ch = as.logical(special_ch),
        special_mm = as.logical(special_mm),
        store = as.factor(store),
        store_id = as.factor(store_id))

summary(orange_juice)
# 653 purchases of CH, 417 purchases of MM
```

```{r}
#alias(not sure what I want to put in here ~ ., data = orange_juice)
# 1. price difference columns (list_price_diff and (sale)price_diff) can probably 
# be derived from the other price columns
# 2. store and store id are the same really. and store7 can be worked out from either
# (i didn't make store_id a factor so going back to do that)

alias(purchase ~ ., data = orange_juice)

```
```{r}
orange_juice <- orange_juice %>% 
  select(-c("store7", "store", "price_diff", "list_price_diff"))
```

```{r}
# orange_juice %>% 
# ggpairs() # had previously changed weekof_purchase to a factor but it has too many levels... so i changed it back

#^ this is too messy to see, so split it up, i guess by half

orange_juice %>% 
  select(c("purchase", "weekof_purchase", "store_id", "price_ch",
           "price_mm", "disc_ch", "disc_mm")) %>% 
  ggpairs()
```
^ `store_id`, `disc_ch`, `disc_mm`, and `price_mm` all seem interesting...  


```{r}
orange_juice %>% 
  select(c("purchase", "special_ch", "special_mm", "loyal_ch", 
           "sale_price_mm", "sale_price_ch", "pct_disc_mm",
           "pct_disc_ch")) %>% 
  ggpairs()
```
^ argument for all, but especially `loyal_ch` and the percentage discounts






