---
title: "SQL Day 2 Lab and Homework - Solutions"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r, eval=TRUE, include=FALSE, message=FALSE, warning=FALSE}
library(RPostgres)

db = dbConnect(
  Postgres(), 
  user = 'omni_user',
  password = 'niytna5mcwsz124i',
  dbname = 'omni_pool',
  host = 'db-postgresql-lon1-20135-do-user-7581050-0.b.db.ondigitalocean.com',
  port = 25061,
  sslmode = 'require',
  bigint = 'numeric'
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

**Duration - 120 minutes**  

In these practice questions you'll return again to the `omni_pool` database, so re-establish your connection in `DBeaver` if it has lapsed. 

<br>
<div class='emphasis'>
**Submission**

When answering these questions, please save your `SQL` script in `DBeaver`, copy it over to your homework repo and then add, commit and push that for submission tonight. Please add section and question numbers as `SQL` comments, like so:

```{sql, eval=FALSE}
--MVP
--Q1
```
</div>
<br>

# MVP 

<br>

**Question 1.**\
<ul>
(a). Find the first name, last name and team name of employees who are members of `teams`.
<br><br>
<details>
<summary>**Hint**</summary>  
We only want employees who are also in the teams table. So which type of join should we use? 
</details>

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  e.first_name, 
  e.last_name, 
  t.name AS team_name
FROM employees AS e INNER JOIN teams AS t
ON e.team_id = t.id;
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
<br>

(b). Find the first name, last name and team name of employees who are members of `teams` and are enrolled in the pension scheme.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  e.first_name, 
  e.last_name, 
  t.name AS team_name
FROM employees AS e INNER JOIN teams AS t
ON e.team_id = t.id
WHERE e.pension_enrol = TRUE;
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
<br>

(c). Find the first name, last name and team name of employees who are members of `teams`, where their team has a charge cost greater than 80. 
<br><br>
<details>
<summary>**Hint**</summary>  
`charge_cost` may be the wrong type to compare with value `80`. Can you find a way to convert it **without changing** the database? 
</details>

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  e.first_name, 
  e.last_name, 
  t.name AS team_name
FROM employees AS e INNER JOIN teams AS t
ON e.team_id = t.id
WHERE CAST(t.charge_cost AS INT) > 80;
```
```{r, echo=FALSE, eval=TRUE}
out
```
  
Can also do this in `PostgreSQL` as:
  
```{sql, connection = db, output.var="out"}
SELECT 
  e.first_name, 
  e.last_name, 
  t.name AS team_name
FROM employees AS e INNER JOIN teams AS t
ON e.team_id = t.id
WHERE t.charge_cost::INT > 80;
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
</ul>
<br>

**Question 2.**\
<ul>
(a). Get a table of all `employees` details, together with their `local_account_no` and `local_sort_code`, if they have them.
<br><br>
<details>
<summary>**Hints**</summary>  
`local_account_no` and `local_sort_code` are fields in `pay_details`, and employee details are held in `employees`, so this query requires a `JOIN`.
<br><br>
What sort of `JOIN` is needed if we want details of **all** employees, even if they don't have stored `local_account_no` and `local_sort_code`?
</details>

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  e.*,
  pd.local_account_no,
  pd.local_sort_code
FROM employees AS e LEFT JOIN pay_details AS pd
ON e.pay_detail_id = pd.id
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
<br>

(b). Amend your query above to also return the `name` of the team that each employee belongs to.
<br><br>
<details>
<summary>**Hint**</summary>  
The name of the team is in the `teams` table, so we will need to do another join.
</details>

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  e.*,
  pd.local_account_no,
  pd.local_sort_code,
  t.name AS team_name
FROM employees AS e LEFT JOIN pay_details AS pd
ON e.pay_detail_id = pd.id
LEFT JOIN teams AS t
ON e.team_id = t.id
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
</ul>
<br>

**Question 3.**\
<ul>
(a). Make a table, which has each employee id along with the team that employee belongs to.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  e.id AS employee_id,
  t.name AS team_name 
FROM employees AS e LEFT JOIN teams as t
ON e.team_id = t.id;
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
<br>

(b). Breakdown the number of employees in each of the teams.
<br><br>
<details>
<summary>**Hint**</summary>   
You will need to add a group by to the table you created above.
</details>

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  COUNT(e.id),
  t.name AS team_name 
FROM employees AS e LEFT JOIN teams as t
ON e.team_id = t.id
GROUP BY t.name;
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
<br>

(c). Order the table above by so that the teams with the least employees come first.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  t.name AS team_name, 
  COUNT(e.id) AS num_employees
FROM teams AS t LEFT JOIN employees AS e
ON e.team_id = t.id
GROUP BY t.name
ORDER BY num_employees ASC
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
</ul>
<br>

**Question 4.**\
<ul>
(a). Create a table with the team id, team name and the count of the number of employees in each team.

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  t.id,  
  t.name,
  COUNT(e.id)
FROM employees AS e
INNER JOIN teams AS t
ON e.team_id = t.id
GROUP BY t.id
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
<br>

(b). The `total_day_charge` of a team is defined as the `charge_cost` of the team multiplied by the number of `employees` in the team. Calculate the `total_day_charge` for each team.
<br><br>
<details>
<summary>**Hint**</summary> 
If you `GROUP BY teams.id`, because it's the primary key, you can `SELECT` any other column of `teams` that you want (this is an exception to the rule that normally you can only `SELECT` a column that you `GROUP BY`).
</details>

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var="out"}
SELECT 
  t.name,
  COUNT(e.id) * CAST(t.charge_cost AS INT) AS total_day_charge
FROM employees AS e
INNER JOIN teams AS t
ON e.team_id = t.id
GROUP BY t.id
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
<br>

(c). How would you amend your query from above to show only those `teams` with a `total_day_charge` greater than 5000?

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var = "out"}
SELECT 
  t.name,
  COUNT(e.id) * CAST(t.charge_cost AS INT) AS total_day_charge
FROM employees AS e
INNER JOIN teams AS t
ON e.team_id = t.id
GROUP BY t.id
HAVING COUNT(e.id) * CAST(t.charge_cost AS INT) > 5000
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>
</ul>
<br>

# Extension 

<br>

**Question 5.**\
How many of the employees serve on one or more committees?  
<br><br>
<details>
<summary>**Hints**</summary> 
All of the details of membership of committees is held in a single table: `employees_committees`, so this doesn't require a join.
<br><br>
Some employees may serve in multiple committees. Can you find the number of distinct employees who serve? [**Extra hint** - do some research on the `DISTINCT()` function].
</details>

<details>
<summary>**Answer**</summary>
```{sql, connection = db, output.var = "out"}
SELECT 
  COUNT(DISTINCT(employee_id)) AS num_employees_on_committees
FROM employees_committees
```
```{r, echo=FALSE, eval=TRUE}
out
```
</details>

<br>

**Question 6.**\
How many of the employees do not serve on a committee?  
<br><br>
<details>
<summary>**Hints**</summary>   
This requires joining over only two tables
<br><br>
Could you use a join and find rows without a match in the join?
</details>

<details>
<summary>**Answer**</summary>  
Here is the 'anti-join' way of doing it:
  
```{sql, connection = db, output.var = "out"}
SELECT 
  COUNT(*) AS num_not_in_committees
FROM employees e
LEFT JOIN employees_committees ec
ON e.id = ec.employee_id 
WHERE ec.employee_id IS NULL
```
```{r, echo=FALSE, eval=TRUE}
out
```
  
Here is a method using two subqueries:
  
```{sql, connection = db, output.var = "out"}
SELECT 
  (SELECT COUNT(id) FROM employees) -
  (SELECT COUNT(DISTINCT(employee_id)) FROM employees_committees)
    AS num_not_in_committees
```
```{r, echo=FALSE, eval=TRUE}
out
```

```{r, echo=FALSE, eval=TRUE}
dbDisconnect(db)
```
</details>
