---
title: "ggplot homework Week_5 Day_1"
output: html_notebook
date: 04/04/2022
---
```{r}
library(tidyverse)
```

```{r}
competitors <- CodeClanData::qb_competitors
device_data <- CodeClanData::qb_device_data
monthly_sales <- CodeClanData::qb_monthly_sales
revenue_breakdown <- CodeClanData::qb_revenue_breakdown
```

___

# 2 MVP

## 2.1 Part 1 

**Question 1.**
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart 
showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(revenue_breakdown) +
  geom_bar(aes(x = Year, y = Revenue, fill = Product), stat = "identity") +
  labs(title = "Sources of revenue in 2018 and 2019")
  
```


**Questino 2.**
Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit 
Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will 
need to use filter() or subset() (base R).

```{r}
kwikbit_unit1_sales <- monthly_sales %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales")

ggplot(kwikbit_unit1_sales) +
  geom_line(aes(x = Date, y = Amount)) +
  labs(x = "Date",
       y = "Total Sales", 
       title = "Monthly sales of the Kwikbit Unit 1 over the last year")
```


**Question 3.**
Make a line chart showing monthly revenue and costs over the last year. This 
data is also in qb_monthly_sales.

```{r}
monthly_revenue_and_costs <- monthly_sales %>% 
  filter(Cashflow != "Kwikbit Unit 1 Sales")

ggplot(monthly_revenue_and_costs) +
  geom_line(aes(x = Date, y = Amount, group = Cashflow, colour = Cashflow)) +
  labs(x = "Date", 
       y = "Total Amount",
       title = "Monthly Revenue Vs Monthly Costs")
```

**Question 4.**
Show annual sales of personal fitness trackers over the last 5 years broken down 
by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
ggplot(competitors) +
  geom_area(aes(x = Year, y = Revenue, fill = Company, colour = Company), alpha = 0.7, size = 0.5, colour = "black") +
  theme(axis.text.y = element_text(angle = 45, hjust = 1)) +
  labs(x = "Year",
       y = "Revenue", 
       title = "Annual sales of competitors personal fitness trackers")
# why does it change revenue values to e.g. "6e+06"? Top revenue
# is 2,216,000 - is it a space issue? Doesn't make a differnce if # I angle it... 
```


**Question 5.**
Now show the sales from the four competitors as a line graph. Include an extra 
layer that shows the data points used to make the lines.


```{r}
ggplot(competitors) +
  geom_line(aes(x = Year, y = Revenue, group = Company, colour = Company)) +
  geom_point(aes(x = Year, y = Revenue, colour = Company)) +
  labs(x = "Year", 
       y = "Total Revenue", 
       title = "Competitors Sales")
```


**Question 6.**
Now the company wants to compare the number of steps that their device counts 
vs. their competitors. Make a line graph of the number of steps throughout time,
and use faceting to compare between companies and people. The data for this is 
in qb_device_data.

**homework answer - my attempts below, if I had combined a few bits of what** 
**I'd tried then I might have gotten there**
```{r}
device_data <- 
mutate(device_data, decimal_hour = hours + (mins/60))

ggplot(device_data) +
  aes(x = decimal_hour, y = counts) +
  geom_line() + 
  facet_grid(id ~ device)
```



Attempt 1 (more below) - this isn't right...?
Struggling with what to put on the x axis especially.
```{r}
 ggplot(device_data) +
  aes(x = hours, y = counts, group = device, colour = id) +
  geom_line() +
  facet_wrap(device ~ id)
```

Attempt 2 - this isn't right...? the lines change colour whichever way round I 
do it.
```{r}
ggplot(device_data) +
  aes(x = counts, y = hours, group = device, colour = id) +
  geom_line() +
  facet_wrap(device ~ id)
```

Attempt 3. Not correct - I surely want more lines than this 
```{r}
total_steps_and_time <- device_data %>% 
  group_by(id, device) %>% 
  mutate(total_steps = sum(counts, na.rm = TRUE), 
            total_time = sum(hours, na.rm = TRUE) + sum(mins/60, na.rm = TRUE))

ggplot(total_steps_and_time) +
  aes(x = id, y = total_steps, group = device, colour = device) +
  geom_line() #+
  # facet_wrap(device ~ id)
```




```{r}
ggplot(device_data) +
  geom_line(
    aes(x = hours, y = counts, group = id, colour = device))
```






