---
title: "Week 6 - Day 1- Probability Homework"
output: html_notebook
date: 11/04/2022
---

# 1.1 The Titanic

```{r}
library(prob)
library(gtools)
library(tidyverse)
```

```{r}
titanic_passengers <- read_csv("5_quiz_and_homework/data/titanic.csv")
titanic_passengers
```
```{r}
titanic_passengers %>% 
  distinct(class)

titanic_crew_or_pax <- titanic_passengers %>% 
  mutate(crew = class %in% c("engineering crew", 
                           "victualling crew", 
                           "restaurant staff", 
                           "deck crew"), .before = 4)
titanic_crew_or_pax
```

Some time after the disaster a journalist wrote that crew members were more 
likely to have survived the sinking than passengers. This was disputed by a 
descendant of one of the surviving crew members, who suggested that the opposite 
was true, i.e. passengers were more likely to have survived than crew.

**Question**
Calculate the conditional probabilities prob(survived∣crew) and
prob(survived∣passenger) to settle this dispute.

*I was wrong - Class answer is below mine*
*I think my contingency table is correct though*
```{r}
prob_of_survival <- titanic_crew_or_pax %>% 
  tabyl(crew, survived) %>% 
  adorn_totals(where = c("row", "col")) %>%  
  adorn_percentages(denominator = "all") %>% 
  adorn_title(col_name = "Class of 'crew'") 
prob_of_survival
# the probability of being crew AND surviving = 0.0956
# the probability of being a passenger AND surviving = 0.2266

prob_crew_survival = titanic_crew_or_pax %>%
  filter(crew == TRUE & survived == "yes") %>%
  summarise(prob = n() / nrow(titanic_crew_or_pax))
prob_crew_survival
# answer = 0.0956

prob_pax_survival = titanic_crew_or_pax %>%
  filter(crew == FALSE & survived == "yes") %>%
  summarise(prob = n() / nrow(titanic_crew_or_pax))
prob_pax_survival
# answer = 0.2266
```
The probability of being a passenger and surviving was higher (22.7%) than that 
of being a member of staff and surviving (9.6%)


*this is the correct answer from class* 
```{r}
titanic_probs <- titanic %>%
  group_by(crew) %>%
  summarise(prob = sum(survived == "yes") / n()) 

titanic_probs
# Crew = 0.24, Passenger = 0.38

#So passengers were 1.6 times more likely to have #survived the disaster than crew.
```









# 1.2 Probability Consolidation Questions

**Question 1.**
If I roll a six-sided dice once, what is the probability of getting a number 
less than 5?

*Answer* 
By hand: 
2/3 which is 0.66.

With R:
```{r}
roll_dice_once <- rolldie(1) 

roll_dice_once %>% 
  mutate(event = X1 < 5) %>% 
  summarise(sum(event), mean(event))
#answer = 0.66667
```




**Question 2.**
How many outcomes are there in the sample space of an experiment in which we 
toss three dice?

*Answer*
```{r}
three_dice <- rolldie(3)

three_dice %>% 
  summarise(size_of_sample = n())

# there are 216 outcomes
```




**Question 3.**
In an experiment where we roll four dice:

How many outcomes are in the event where the total of the four dice equals 5, 6 
or 7)?
what is the probability of this event?
what assumptions do we apply in answering part ii?
Hint
For part i, rowSums(.) may help you setup up the necessary random variable total 
(or you can use pivot_longer())

*Answer*
```{r}
four_dice <- rolldie(4) %>% 
  mutate(four_dice_total = rowSums(.))

four_dice_5_6or7 <- four_dice %>% 
  filter(four_dice_total >= 5 & four_dice_total <= 7)
  
four_dice_5_6or7 %>% 
  summarise(size_of_sample = n())
# Sample size for four dice being rolled where the total of the four dice is 
# 5, 6, or 7, is 34 
  

prob_5_6_or_7 <- four_dice_5_6or7 %>%
  summarise(prob = n() / nrow(four_dice))
prob_5_6_or_7

# the probability of rolling 4 dice and seeing a total of 5, 6, or 7, is 0.03 
#(0.026)

# Here, the assumption I have applied is that order matters and so repeats are 
# allowed. I.e. there is 2, 1, 1, 1 and also 1, 2, 1, 1 etc. 
```




**Question 4.**
Five chairs are placed in a row in a room, along with five people of differing 
height. The people sit down randomly on the chairs. What is the probability that 
they sit in order of decreasing height, left to right?

Hint
This is an arrangement where order matters. Of all the possible arrangements, 
how many are correct?

*Answer*
```{r}
chairs <- 1:5
people <- c("tallest", "tall", "average", "short", "shortest")
tibble(chairs = chairs, height = people)

permutations(n = 5, r = 5, v = people,
             repeats.allowed = FALSE)
# there are 120 permutations so 1/120
# probability = 0.008

```



