---
title: "Week_5 Day_2 Homework_mahri"
output: html_notebook
---
## Introduction to ggplot2 homework

**Question 1.**
Load in the libraries you need, and look at the backpack data. Take note of 
variable names and dimensions.

```{r}
library(tidyverse)
library(ggthemes)
```

```{r}
backpack <- CodeClanData::backpack
backpack
dim(backpack)
glimpse(backpack)
names(backpack)
```

The dataset "backpack" has 100 rows and 9 columns with column names: 
* backpack_weight - ranging from 2 - 35
* body_weight - ranging from 105 - 270
* ratio 
* back_problems - a numeric column where 0 = "no back problems" and 1 = "back 
problems"
* major - the individual's University Major
* year - ranging from 0 - 6
* sex - "male" or "female"
* status - "U" or "G"
* units - ranging from 0 - 20



**Question 2.**
What is the distribution of backpack weights that students are carrying? Choose 
an appropriate graph to show the distribution.

**Answer** 
Hopefully the weight of backpacks is not measured in kg's as there is one 
outlying student carrying a 35kg bag. Most of the students are carring bags of
between 5 and 15kgs with the lightest bag weighing 2kg.

```{r}
backpack_weights <- backpack %>% 
  group_by(backpack_weight) %>% 
  summarise(count = n())

ggplot(backpack_weights) +
  aes(x = backpack_weight, y = count) +
  geom_col() +
  scale_x_continuous(breaks = seq(2, 36, 2)) +
  scale_y_continuous(breaks = seq(2, 14, 2)) +
  labs(x = "Weight of backpack (kgs)", 
       y = "Number of students", 
       title = "Distribution of backpack weights being carried by students")
```




**Question 3.**
Are there differences in backpack weight between sex?

**Answer** 
find the mean back pack weight?
```{r}

ggplot(backpack) +
  aes(x = backpack_weight, y = sex) +
  geom_line()
```


**Question 4.**
What is the relationship between body weight and backpack weight?

**Answer**
There is not a very clear relationship between body weight and backback weight, 
however there might be a slight arguement for an increase in one relating to an 
increase in the other. 

```{r}
backpack
ggplot(backpack)+
  aes(x = body_weight, y = backpack_weight) +
  geom_point() +
  labs(x = "Body weight (lbs)", 
       y = "Backpack weight (kgs)", 
       title = "The relationship between body weight and backpack weight")

```


**Question 5.**
Is this relationship different between men and women? Choose an appropriate 
plot and interpret your findings.

**Answer**
Men tend to be heavier than women however I could not argue that the 
relationship between body weight and backpack weight is different between sexes.

WORK OUT MEANS TO CHECK THIS 

```{r}
backpack
ggplot(backpack)+
  aes(x = body_weight, y = backpack_weight, colour = sex) +
  geom_point() +
  labs(x = "Body weight (lbs)", 
       y = "Backpack weight (kgs)",
       colour = "Sex",
       title = "The relationship between body weight and backpack weight")

```


**Question 6.**
What proportion of students report back pain?

**Answer** 
32% (just less than 1/3) of students report to have back problems 
and 68% (just more than 2/3) report none.
```{r}
backpack %>% 
  group_by(back_problems) %>% 
  summarise(count = n())
  
ggplot(backpack) +
  aes(x = back_problems) +
  geom_bar() +
  labs(x = "No back problems - Vs - Back problems",
       y = "Number of students",
       title = "Proportion of students reporting backpain")



# got to work out how to clean up the below so that it works...
ggplot(backpack) +
  aes(x = back_problems, fill = back_problems == "1") +
  geom_bar() +
  scale_fill_manual(values = c("green", "red")) +
  labs(x = "No back problems  vs back problems",
       y = "Number of students",
       title = "Proportion of students reporting backpain") +
  geom_text(aes(label = count),                  
           color = "black", size = 4, hjust = 0)
  
  
  ggplot(total_sales_anno) +
  aes(x = branch, y = sales, fill = branch == "Leeds") +
  geom_col() +
  scale_fill_manual(values = c('grey40', 'red'), 
                    guide = "none") +
  coord_flip() +
    geom_text(aes(label = sales),                  
           color = "red", size = 4, hjust = 0) 
```


**Question 7.**
Is this proportion different between men and women?

**Answer**
Yes but I can't yet think how to do the sum properly 
```{r}
backpack %>% 
  group_by(sex, back_problems) %>% 
  summarise(count = n())
```

**Question 8.**
Now you’ve investigated your data separately, create plots to describe the 
relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

```{r}

ggplot(backpack) + 
  aes(x = body_weight, y = backpack_weight, 
      group = sex) +
  geom_line() +
  facet_wrap(back_problems~sex)


```



